---
title: "exploring-summary-stats"
author: "K.L. Beck"
date: "2022-10-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(arrow) # read parquet
library(cluster)
library(dplyr)
library(ggplot2)
library(tidyverse)

options(scipen = 10)
```

```{r joinGOec}

ec_go = read_parquet("~/Work/OMXWare-Local/Virulence/part-00000-24dd3420-9bdc-4922-9e7d-e643183da818-c000.snappy.parquet")

ipr_go = read_parquet("~/Work/OMXWare-Local/Virulence/part-00000-094914b3-df54-45e0-a968-ccdc7d34c2b3-c000.snappy.parquet")

go_table = read_parquet("~/Work/OMXWare-Local/Virulence/part-00000-d0b15c47-52ef-49b9-8dd8-83f7e432f458-c000.snappy.parquet")

domain_architectures_ipr_typed = read.csv("~/Work/Code/OMXW/virulence/outputs/domain_architecture.csv", stringsAsFactors = FALSE)

da2_ipr_long = domain_architectures_ipr_typed %>% dplyr::select(DOMAIN_ARCHITECTURE_UID_KEY, DOMAIN_ARCHITECTURE) %>% tidyr::separate_rows(DOMAIN_ARCHITECTURE, sep = "~", convert = FALSE)
colnames(da2_ipr_long)[2] = "IPR_ACCESSION"

da2_ipr = domain_architectures_ipr_typed %>% dplyr::select(DOMAIN_ARCHITECTURE_UID_KEY, DOMAIN_ARCHITECTURE)

```


## Questions of Note

### Q1 Which virulence discoveries standout by count (top 25 across all organisms)? What are the protein names associated with these?

```{r standoutDiscoveriesProtNames}
standout_da_counts = read.csv("~/Work/Code/OMXW/virulence/outputs/standout_discoveries.csv", stringsAsFactors = FALSE)
standout_da_prot_names = read.csv("~/Work/Code/OMXW/virulence/outputs/standout_discoveries_protein.csv", stringsAsFactors = FALSE)

summary(standout_da_counts$COUNT)

ggplot(standout_da_counts, aes(COUNT)) +
  geom_histogram(bins = 100) +
  xlab("Domain Architecture Number of Observations")

ggplot(standout_da_counts, aes(x="Discovery", y=COUNT)) +
  geom_violin(width = 1.5) +
  geom_boxplot(width = 0.1, color="grey", alpha=0.2) +
  scale_y_log10() +
  ylab("Domain Architecture Number of Observations")
# TODO compare this with copivots?

top_discoveries =  standout_da_counts %>% dplyr::slice_max(COUNT, n= 20)

standouts = merge(standout_da_prot_names, standout_da_counts)


# Q1
top_discoveries_named = standouts[which(standouts$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY %in% top_discoveries$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY), ]

# how many protein names are there per top DAs?
table(top_discoveries_named$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY)
mean(table(top_discoveries_named$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY))
range(table(top_discoveries_named$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY))

# TODO are there names that are predominant for a DA (to help resolve synonyms)? ask Ed
# add count per protein to each row in this table standout_discoveries_protein.csv


# frequency of protein names for top das
prot_name_counts_to_discovered_da = data.frame(table(top_discoveries_named$PROTEIN_FULLNAME))

# these are the enriched protein names that are in high occurring discovered DAs
# and this protein name occurs in more than 1 top discovery
enriched_protein_names  = prot_name_counts_to_discovered_da[which(prot_name_counts_to_discovered_da$Freq > 1 & !(prot_name_counts_to_discovered_da$Var1 %in% c("hypothetical protein", "putative protein"))), ]

enriched_protein_names_expanded = merge(top_discoveries_named, enriched_protein_names, by.y = "Var1", by.x = "PROTEIN_FULLNAME")

# good intermediate plot, but really should show cumulative counts
ggplot(enriched_protein_names_expanded, aes(x=Freq, y=COUNT, label=PROTEIN_FULLNAME, color = PROTEIN_FULLNAME)) +
  geom_point() +
  geom_text(hjust = 0, vjust=0) +
  xlab("Present in N Discovered Domain Architectures") +
  ylab("Observations as Discovered Domain Architecture")

# Q1 - manuscript version
ggplot(enriched_protein_names_expanded, aes(x=COUNT, y=PROTEIN_FULLNAME, fill = NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY)) +
  geom_bar(position = "stack", stat="identity") +
  scale_fill_brewer(palette = "Paired") +
  xlab("DOMAIN ARCHITECTURE OBSERVATIONS") +
  ggtitle("Enriched protein names for top discovered domain architectures")


# merge top discoveries with GO and with EC
top_discoveries_func = merge(merge(merge(merge(top_discoveries, da2_ipr_long, by.x = "NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY", by.y = "DOMAIN_ARCHITECTURE_UID_KEY"), ipr_go, all.x = TRUE), ec_go, all.x = TRUE), go_table, all.x = TRUE) 
top_discoveries_func$GO_LABEL = paste(top_discoveries_func$GO_TERM, top_discoveries_func$DESCRIPTION, sep="|")

# drop missing GO terms for viz
top_discoveries_func_GO_complete = top_discoveries_func[!is.na(top_discoveries_func$GO_TERM), ]

# Q1 expanded with GO - manuscript version
ggplot(top_discoveries_func_GO_complete, aes(x=COUNT, y=GO_LABEL, fill = NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY)) +
  geom_bar(position = "stack", stat="identity") +
  xlab("DOMAIN ARCHITECTURE OBSERVATIONS") +
  ggtitle("GO Terms for top discovered domain architectures")

# three DAs don't have GO terms - true in latest release
# only 2 have ec codes

# try look up here to extract more ec codes, if time
# https://ftp.ebi.ac.uk/pub/databases/interpro/current_release/interpro.xml.gz


```


### Q1a For the high occurring discoveries, do they occur with a diversity of virulence factors (pivots)?
    - Q1b For the newly discovered proteins occurring with low diversity pivots (consistently occurring D-P pairs), is
      there a diversity of function (IPR codes) indicating functional dependence?
        - Q1b1 For the neighbors, can you identify that the neighbors are regulatory in nature (and expected)? Or
          related to excretion?

```{r topDiscoveryPivots}

pivot_discovery = read.csv("~/Work/Code/OMXW/virulence/outputs/pivot_discovery_frequency.csv", stringsAsFactors = FALSE)

pivot_div_discovery = pivot_discovery[which(pivot_discovery$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY %in% unique(top_discoveries$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY)), ]


# Q1a
# this is the number of distinct pivot pairings for each top discovered da
table(pivot_div_discovery$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY)
mean(table(pivot_div_discovery$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY))
range(table(pivot_div_discovery$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY))
# note, this doesn't factor in which of the pairings are most common since COUNT isn't considered here

# total number of pivots neighboring standout discovered DAs
length(unique(pivot_div_discovery$PIVOT_DOMAIN_ARCHITECTURE_UID_KEY))
length(unique(pivot_div_discovery$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY)) # standout discovered DAs, sanity check

# pivot frequency for top standout discoveries
pivots_for_standout_discoveries = data.frame(table(pivot_div_discovery$PIVOT_DOMAIN_ARCHITECTURE_UID_KEY))
# there are pivots that occur with all top 20 das
# what are those?

ggplot(pivots_for_standout_discoveries, aes(Freq)) +
  geom_histogram(binwidth = 1, col="white") +
  stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1) +
  xlab("Pivot DAs occurrence with standout discovered DAs (max 20)") +
  ggtitle("Histogram of pivot occurrence for standout discovered DAs")

pivot_uids_common_in_standout_discoveries = pivots_for_standout_discoveries$Var1[which(pivots_for_standout_discoveries$Freq > 15)]
write.csv(pivot_uids_common_in_standout_discoveries, "~/Work/Code/OMXW/virulence/outputs/pivot_uids_for_standout_discoveries.csv", quote = FALSE, row.names = FALSE)

# what are the pivot protein names associated with those standout discoveries?
# TODO next
pivot_prot_names = read.csv("~/Work/Code/OMXW/virulence/outputs/standout_pivot_proteins (1).csv", stringsAsFactors = FALSE)
range(table(pivot_prot_names$DOMAIN_ARCHITECTURE_UID_KEY)) # how many protein names per uid?
mean(table(pivot_prot_names$DOMAIN_ARCHITECTURE_UID_KEY))


pivot_prot_names_top = pivot_prot_names %>% group_by(DOMAIN_ARCHITECTURE_UID_KEY) %>% dplyr::slice_max(COUNT, n=1)
pivot_prot_names_top_no_hypo = pivot_prot_names_top[which(pivot_prot_names_top$PROTEIN_FULLNAME != "hypothetical protein"), ]
# pulled top 20 protein names by occurrence to include in slides


test = data.frame(animal = c("cat", "dog", "fish", "dog"),
                  trait = c("furry", "noisy", "wet","furry"),
                  presence = c(3,1,1,4))

test = data.frame(animal = c("cat", "dog", "fish", "dog"),
                  trait = c("furry", "noisy", "wet","furry"),
                  presence = 1)



ggplot(test, aes(animal, trait, fill = presence)) +
  geom_tile() 

tmp = spread(test, animal, presence, fill = 0)
row.names(tmp) = tmp$trait
tmp$trait = NULL
heatmap(as.matrix(tmp), scale = "none")

ggplot(test, aes(animal, trait, group = 1:nrow(test))) +
  geom_point()

ggplot(pivot_div_discovery, aes(PIVOT_DOMAIN_ARCHITECTURE_UID_KEY, NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY)) +
  geom_point()

# pivot-discovery (all pairings) with frequency
pivot_div_discovery_mat = spread(pivot_div_discovery, NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY, COUNT, fill = 0)
row.names(pivot_div_discovery_mat) = pivot_div_discovery_mat$PIVOT_DOMAIN_ARCHITECTURE_UID_KEY
pivot_div_discovery_mat$PIVOT_DOMAIN_ARCHITECTURE_UID_KEY = NULL
heatmap(as.matrix(pivot_div_discovery_mat), scale = "none") # no color normalization
heatmap(as.matrix(pivot_div_discovery_mat))

# TODO cleanup heatmap with color frequency ^

# Q1a
heatmap(as.matrix(pivot_div_discovery_binary_mat), 
        scale = "none",
        margins = c(10,10),
        xlab = "Standout discovered DAs",
        cexCol = 0.5,
        ylab = "Pivot DAs",
        main = "Observation of Pivot DAs with Standout Discovered DAs") # no color normalization



# just pivot protein presence/absence with standout discoveries
pivot_div_discovery_binary = pivot_div_discovery
pivot_div_discovery_binary$COUNT = 1
pivot_div_discovery_binary_mat = spread(pivot_div_discovery_binary, NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY, COUNT, fill = 0)
row.names(pivot_div_discovery_binary_mat) = pivot_div_discovery_binary_mat$PIVOT_DOMAIN_ARCHITECTURE_UID_KEY
pivot_div_discovery_binary_mat$PIVOT_DOMAIN_ARCHITECTURE_UID_KEY = NULL

# Q1a
heatmap(as.matrix(pivot_div_discovery_binary_mat), 
        scale = "none",
        margins = c(10,10),
        xlab = "Standout discovered DAs",
        cexCol = 0.5,
        ylab = "Pivot DAs",
        main = "Observation of Pivot DAs with Standout Discovered DAs") # no color normalization

```

### Q2 For all discovered domain architectures, what has the highest count for each genera (top 50)? Map those to GO terms
  for enriched pathways. What are the protein names associated with these?
    - Q2a Are there discovered DAs unique to genera?
    - Q2b Are there discovered DAs that are present across multiple genera?
    
### Q3 Which IPR codes occur in the most discovered DAs?
TODO

### Q4 What are the most frequent pairs of pivot and co-pivots?
```{r p_co_pairing}
pivot_copivot_freq = read.csv("~/Work/Code/OMXW/virulence/outputs/pivot_copivot_frequency.csv", stringsAsFactors = FALSE)

length(unique(pivot_copivot_freq$PIVOT_DOMAIN_ARCHITECTURE_UID_KEY))
length(unique(pivot_copivot_freq$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY))
summary(pivot_copivot_freq$COUNT)

# remove self-comparisons
pivot_copivot_freq_no_self = pivot_copivot_freq[-which(pivot_copivot_freq$PIVOT_DOMAIN_ARCHITECTURE_UID_KEY == pivot_copivot_freq$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY), ]

# what do we do with mirrors? the table gives a full matrix and we only want the half matrix so counts aren't inflated
# e.g. A - B, 5
#      B - A, 5
# only count each pair either A - B or B - A once

pivot_copivot_freq_no_self$ORDERED_DA_STR = ""
pivot_copivot_freq_no_self$ORDERED_DA_STR = 
  sapply(1:nrow(pivot_copivot_freq_no_self), function(n)
    paste(sort(c(pivot_copivot_freq_no_self$PIVOT_DOMAIN_ARCHITECTURE_UID_KEY[n], pivot_copivot_freq_no_self$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY[n])), collapse = "~")
    )

# verify that the per group count is the same regardless for A - B and B - A
n_distinct_by_DA_pair = pivot_copivot_freq_no_self %>% group_by(ORDERED_DA_STR) %>% summarize(n_values = n_distinct(COUNT))
all(n_distinct_by_DA_pair$n_values == 1) # each pivot co-pivot pair should have 1 count value

# deduplicate to get half matrix
pivot_copivot_freq_clean = pivot_copivot_freq_no_self[!duplicated(pivot_copivot_freq_no_self$ORDERED_DA_STR), ]

# cleanup
rm(pivot_copivot_freq_no_self)


# is there a pivot-copivot pair that always appear together?

# compute counts as relative value since pivot occurrence is a function of what's been sequenced
# calculate percent co-pivot appearance with pivot
# e.g. co-pivots that always appear with pivots would be 100%
# pivots that occur with a diversity of co-pivots would have many co-pivots and those percentages would be lower

pivot_freq_sum = pivot_copivot_freq_clean %>% group_by(PIVOT_DOMAIN_ARCHITECTURE_UID_KEY) %>% summarize(PIVOT_SUM = sum(COUNT))

n_genomes = 206575
pivot_copivot_relative = merge(pivot_copivot_freq_clean, pivot_freq_sum)
pivot_copivot_relative$PERCENT_COPIVOT_OCCURRENCE = pivot_copivot_relative$COUNT / pivot_copivot_relative$PIVOT_SUM * 100
pivot_copivot_relative$PIVOT_PERCENT_TOTAL_GENOMES = pivot_copivot_relative$PIVOT_SUM / n_genomes * 100

summary(pivot_copivot_relative$PERCENT_COPIVOT_OCCURRENCE)
summary(pivot_copivot_relative$PIVOT_PERCENT_TOTAL_GENOMES) # can be greater than 100% if domain architecture occurs multiple times in genomes

length(which(pivot_copivot_relative$PERCENT_COPIVOT_OCCURRENCE > 99))

# does the pure co-pivot percentage just inflate the low ones?
ggplot(pivot_copivot_relative, aes(y=COUNT, x=PIVOT_SUM, color = PERCENT_COPIVOT_OCCURRENCE)) +
  geom_point() +
  scale_color_continuous() +
  xlab("Total observations of pivot DA") +
  xlim(c(0,2500000)) +
  ylim(c(0,2500000)) +
  ylab("Count of co-pivot observations per pivot")

# sanity check: what are some of these very high pivot occurrences?
da2_ipr_long[which(da2_ipr_long$DOMAIN_ARCHITECTURE_UID_KEY == "abf0fb6ee6c6d9aae6872e0e026a7e13"), ]
# https://www.ebi.ac.uk/interpro/entry/InterPro/IPR003439/
# ABC transporter-like, ATP-binding domain
# highly conserved across bacteria, 50 different protein subfamilies
# present in multiple places in genome
# sanity checked

# maybe consider this for supplement? all points
ggplot(pivot_copivot_relative, aes(x=PERCENT_COPIVOT_OCCURRENCE, y=PIVOT_PERCENT_TOTAL_GENOMES, color = PERCENT_COPIVOT_OCCURRENCE)) +
  geom_point() +
  scale_color_continuous() +
  xlab("Co-pivot percent occurrence per pivot") +
  ylab("Pivot percent per total genomes") +
  geom_hline(yintercept = 100)

pivot_copivot_relative_simple = pivot_copivot_relative %>% select(PERCENT_COPIVOT_OCCURRENCE, PIVOT_PERCENT_TOTAL_GENOMES)

set.seed(527)
clustered_pivot_copivot_relative = kmeans(pivot_copivot_relative_simple, centers = 15)

# TODO what are the right numerical cutoffs for commonly present pivot co-pivot pairs? can k be determined without eyeballing data
# work in progress - commenting out, hitting memory cap in R studio
# consider moving to python
# silhouette_score <- function(k){
#   km <- kmeans(pivot_copivot_relative_simple, centers = k, nstart=25)
#   ss <- silhouette(km$cluster, dist(pivot_copivot_relative_simple))
#   mean(ss[, 3])
# }
# k <- 2:5
# 
# subsample_indices = sample(1:nrow(pivot_copivot_relative_simple), size = 50000)
# 
# km <- kmeans(pivot_copivot_relative_simple, centers = 3)
# ss <- silhouette(km$cluster[subsample_indices], dist(pivot_copivot_relative_simple[subsample_indices,]))
# mean(ss[, 3])
# 
# avg_sil <- sapply(k, silhouette_score)
# plot(k, type='b', avg_sil, xlab='Number of clusters', ylab='Average Silhouette Scores', frame=FALSE)

# cut to single copy proteins
ggplot(pivot_copivot_relative, aes(x=PERCENT_COPIVOT_OCCURRENCE, y=PIVOT_PERCENT_TOTAL_GENOMES, color = as.factor(clustered_pivot_copivot_relative$cluster))) +
  geom_point() +
  xlab("Co-pivot percent occurrence per pivot") +
  ylab("Pivot percent per total genomes") +
  ylim(c(0,100)) +
  labs(colour = "k-means cluster")
  # geom_hline(yintercept = 75, color = "red") +
  # geom_vline(xintercept = 75, color = "red")

# all pairings - maybe for supplement if we don't use plot above
ggplot(pivot_copivot_relative, aes(x=PERCENT_COPIVOT_OCCURRENCE, y=PIVOT_PERCENT_TOTAL_GENOMES, color = as.factor(clustered_pivot_copivot_relative$cluster))) +
  geom_point() +
  xlab("Co-pivot percent occurrence per pivot") +
  ylab("Pivot percent per total genomes") +
  labs(colour = "k-means cluster")

# number of pivot co-pivot pairs
length(which(pivot_copivot_relative$PIVOT_PERCENT_TOTAL_GENOMES > 50 & pivot_copivot_relative$PERCENT_COPIVOT_OCCURRENCE > 50))

# TODO-next what are the pivot and co-pivot protein names that are co-occurring in the mid-green cluster? 
# e.g. those high in genome corpus and high in co-occurring rate
# these are bacteria-wide conserved virulence features
pivot_copivot_relative_w_cluster = cbind(pivot_copivot_relative, "KMEANS_CLUSTER" = clustered_pivot_copivot_relative$cluster)
pivot_copivot_relative_w_enriched_cluster = pivot_copivot_relative_w_cluster[which(pivot_copivot_relative_w_cluster$KMEANS_CLUSTER == 5), ]
# this is the name table for protein looks up ^

conserved_pairings = pivot_copivot_relative_w_cluster


# next calculate n_distinct per pivot to get diversity



```
### Q5 What are the most frequent pairs of pivot and discoveries?

```{r p_disc_pairing}
# TODO generate comparable k-means clustering plot for pivot-discovery pair with percentage occurrence

pivot_discovered_freq = read.csv("~/Work/Code/OMXW/virulence/outputs/pivot_discovery_frequency.csv", stringsAsFactors = FALSE)

length(unique(pivot_discovered_freq$PIVOT_DOMAIN_ARCHITECTURE_UID_KEY))
length(unique(pivot_discovered_freq$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY))
summary(pivot_discovered_freq$COUNT)

# verify there are no self comparisons
length(which(pivot_discovered_freq$PIVOT_DOMAIN_ARCHITECTURE_UID_KEY == pivot_discovered_freq$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY))

# verify there are no mirrors (like there are with pivot-copivot pairs)
table(unique(pivot_discovered_freq$PIVOT_DOMAIN_ARCHITECTURE_UID_KEY) %in% unique(pivot_discovered_freq$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY))
table(unique(pivot_discovered_freq$NEIGHBOR_DOMAIN_ARCHITECTURE_UID_KEY) %in% unique(pivot_discovered_freq$PIVOT_DOMAIN_ARCHITECTURE_UID_KEY))


# is there a pivot-discovered pair that always appear together?

# compute counts as relative value since pivot occurrence is a function of what's been sequenced
# calculate percent discovered appearance with pivot
# e.g. discoveries that always appear with pivots would be 100%
# pivots that occur with a diversity of discoveries would have many discoveries and those percentages would be lower

pivot_freq_sum_d = pivot_discovered_freq %>% group_by(PIVOT_DOMAIN_ARCHITECTURE_UID_KEY) %>% summarize(PIVOT_SUM = sum(COUNT))
# pivot frequency sums for pivot-discovered pairings

# observation:
# the sum of pivot observations with discoveries is higher than pivot observations with co-pivots
# there are also more discovery uids, e.g. discovery diversity higher than co-pivots
# QQ - how do the counts of discoveries compare to the counts of co-pivots though?
# this is something that would be comparable between sets since n genomes is same for both calculations

pivot_discovered_relative = merge(pivot_discovered_freq, pivot_freq_sum_d)
pivot_discovered_relative$PERCENT_DISCOVERED_OCCURRENCE = pivot_discovered_relative$COUNT / pivot_discovered_relative$PIVOT_SUM * 100
pivot_discovered_relative$PIVOT_PERCENT_TOTAL_GENOMES = pivot_discovered_relative$PIVOT_SUM / n_genomes * 100

summary(pivot_discovered_relative$PERCENT_DISCOVERED_OCCURRENCE)
summary(pivot_discovered_relative$PIVOT_PERCENT_TOTAL_GENOMES) # can be greater than 100% if domain architecture occurs multiple times in genomes

length(which(pivot_discovered_relative$PERCENT_DISCOVERED_OCCURRENCE > 99))

# TODO make sense of the plot below and press on
# TODO start here

# QQ - why are there so many pivots with discoveries? does this mean something biologically?
# is this an artifact of the data processing?
max(pivot_discovered_relative$COUNT)
max(pivot_discovered_relative$PIVOT_SUM)
max(pivot_copivot_relative$COUNT)
max(pivot_copivot_relative$PIVOT_SUM)


# does the pure discovered percentage just inflate the low ones?
ggplot(pivot_discovered_relative, aes(y=COUNT, x=PIVOT_SUM, color = PERCENT_DISCOVERED_OCCURRENCE)) +
  geom_point() +
  scale_color_continuous() +
  xlab("Total observations of pivot DA") +
  xlim(c(0,14000000)) +
  ylim(c(0,14000000)) +
  ylab("Count of discovered observations per pivot")


# maybe consider this for supplement? all points
ggplot(pivot_discovered_relative, aes(x=PERCENT_DISCOVERED_OCCURRENCE, y=PIVOT_PERCENT_TOTAL_GENOMES, color = PERCENT_DISCOVERED_OCCURRENCE)) +
  geom_point() +
  scale_color_continuous() +
  xlab("Discovered percent occurrence per pivot") +
  ylab("Pivot percent per total genomes") +
  geom_hline(yintercept = 100) # single copy pivots

pivot_discovered_relative_simple = pivot_discovered_relative %>% select(PERCENT_DISCOVERED_OCCURRENCE, PIVOT_PERCENT_TOTAL_GENOMES)

set.seed(527)
clustered_pivot_discovered_relative = kmeans(pivot_discovered_relative_simple, centers = 10)

# cut to single copy proteins
ggplot(pivot_discovered_relative, aes(x=PERCENT_DISCOVERED_OCCURRENCE, y=PIVOT_PERCENT_TOTAL_GENOMES, color = as.factor(clustered_pivot_discovered_relative$cluster))) +
  geom_point() +
  xlab("Discovered percent occurrence per pivot") +
  ylab("Pivot percent per total genomes") +
  ylim(c(0,100)) +
  labs(colour = "k-means cluster")
  # geom_hline(yintercept = 75, color = "red") +
  # geom_vline(xintercept = 75, color = "red")

# all pairings - maybe for supplement if we don't use plot above
ggplot(pivot_discovered_relative, aes(x=PERCENT_DISCOVERED_OCCURRENCE, y=PIVOT_PERCENT_TOTAL_GENOMES, color = as.factor(clustered_pivot_discovered_relative$cluster))) +
  geom_point() +
  xlab("Discovered percent occurrence per pivot") +
  ylab("Pivot percent per total genomes") +
  labs(colour = "k-means cluster")

# silhouette coefficient
# -1 to 1 higher better

# observations
# there really aren't any highly conserved discoveries across all organisms
# e.g. > 75% and >75%
# maybe better to look at things by genus
# TODO look at the discoveries that do have high observation with the pivots though
# that may just relate to the top discoveries plots from earlier though

# TODO label points with protein names and GO codes for enriched function + add to paper

```

### Q6 Basic stats by genome of counts: Total virulence factors per genome, Total discoveries per genome, Total co-pivots per genome
```{r perGenomeCounts}
per_genome = read.csv("~/Work/Code/OMXW/virulence/outputs/genome_basic_stats.csv", stringsAsFactors = FALSE)

table(per_genome$NEIGHBOR_TYPE)
length(unique(per_genome$ACCESSION_NUMBER))
per_genome$LONG_NEIGHBOR_TYPE = ""
per_genome$LONG_NEIGHBOR_TYPE[which(per_genome$NEIGHBOR_TYPE == "C")] = "Co-pivot"
per_genome$LONG_NEIGHBOR_TYPE[which(per_genome$NEIGHBOR_TYPE == "D")] = "Discovered"
per_genome$LONG_NEIGHBOR_TYPE[which(per_genome$NEIGHBOR_TYPE == "P")] = "Pivot"


# manuscript version
ggplot(per_genome, aes(x=LONG_NEIGHBOR_TYPE, y=COUNT, fill = LONG_NEIGHBOR_TYPE)) +
  geom_violin(width = 1.5) +
  geom_boxplot(width = 0.1, color="grey", alpha=0.2) +
  scale_fill_viridis_d() + 
  scale_y_log10() +
  theme(legend.position = "none") +
  ylab("Count per genome") +
  xlab("Domain architecture type") +
  ggtitle("Distributions of domain architecture counts per genome")

# plot everything, slow to load
# ggplot(per_genome, aes(x=ACCESSION_NUMBER, y=COUNT, fill = NEIGHBOR_TYPE)) +
#   geom_bar(position = "stack", stat="identity") +
#   theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

per_genome %>% group_by(NEIGHBOR_TYPE) %>% summarize(type_range=range(COUNT))
per_genome %>% group_by(NEIGHBOR_TYPE) %>% summarize(type_mean=mean(COUNT))
per_genome %>% group_by(NEIGHBOR_TYPE) %>% summarize(type_median=median(COUNT))


# TODO if time/worthwile, we could pull the actual occurrence by DA type 
# from the frequency pairing dfs above and do an updated violin of DA loadings basically
# would need to add type to each df and extract pivot sums
# pivot sums could be a little skewed though since a pivot could be counted in each dataframe
# but the discovered occurrence counts and the co-pivot occurrence counts would be accurate


```








